
---
title: "Reproducible Research: Peer Assessment 1" 
author: "Joanna Tyszkiewicz Georgescu"
date: "September 21, 2018"
output:
  html_document: default
  keep md: true
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=TRUE, include=FALSE}
##Packages and libraries
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("mice")
install.packages("timeDate")

library(dplyr)
library(mice)
library(timeDate)
library(ggplot2)
```

##1.Loading and preprocessing the data.
###Download data from: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
```{r echo = TRUE, include=FALSE}
##Load data
setwd("C:\\Users\\Joanna\\Desktop\\R_coursera\\5_2_2\\")
path <- getwd()
path
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, file.path(path, "data.zip"))
unzip("data.zip")
```

##2. What is mean total number of steps taken per day?
###Remove NAs, group by date and sum steps.

```{r echo = TRUE, include=FALSE, results='hide', messaga=FALSE, warning=FALSE}
##Remove NAs
d <- read.csv("activity.csv", header = TRUE,  na.strings = c(" ",NA))
na.omit(d)
```
```{r echo=TRUE, results='hide', include=TRUE}
avg <- na.omit(d %>% group_by(date) %>% summarise(totalSteps = sum(steps)))
```
###Plot histogram of total steps.
```{r echo=TRUE, results='asis', message=FALSE, warning=FALSE}
##Histogram of steps
qplot(avg$totalSteps, geom="histogram",
      main = "Histogram of total number of steps taken", 
      xlab = "Steps", 
      ylab="Frequency",
      fill=I("yellow"), 
      col=I("red"))
```


```{r echo=TRUE, results='FALSE'}
##Mean and median
summary(avg)
```

###Mean: 10765
###Median: 10766 

##3. What is the average daily activity pattern?
###Remove NAs, group by date and calculate average number of steps.
```{r echo=TRUE, results='hide'}
s <- na.omit(d %>% group_by(date) %>% summarise(avgSteps = mean(steps)))
```
###Time series plot.
```{r}
ggplot(s, aes(date, avgSteps)) + geom_point(na.rm=TRUE, color="blue", size=2) + 
  theme(axis.text.x = element_text(angle=90))
```

###5-min interval with max # of steps
```{r ech=TRUE, results='hide'} 
d <- read.csv("activity.csv", header = TRUE,  na.strings = c(" ",NA))
groupedByInterval <-na.omit(d %>% group_by(interval)) %>% summarise(totSteps = sum(steps))
```

```{r echo=TRUE, include=FALSE, results='hide', warning=FALSE, message=FALSE}
head(arrange(groupedByInterval, -totSteps),1)
```

###Interval:835

###totSteps: 10927


##4. Imputing missing values.
```{r  echo=TRUE, include=FALSE, results='hide'}
##Imputing missing values
d <- read.csv("activity.csv", header = TRUE,  na.strings = c(" ",NA))
sapply(d, function(x) sum(is.na(x)))
## structure
str(d)
##missing data patterns
md.pattern(d)
## impute (using CART)
tempData <- mice(d, m=1, method='cart', printFlag=FALSE)
summary(tempData)
cleanData <- complete(tempData,1)
cleanData
summary(cleanData)
##check missing data pattern
md.pattern(cleanData)
```


###Histogram of steps (data with no missing values).

```{r ech=TRUE, include=TRUE, results='asis'}
qplot(cleanData$steps, geom="histogram",
      binwidth = 50,
      main = "Histogram of total number steps taken", 
      xlab = "Steps", 
      ylab="Frequency",
      fill=I("red"), 
      col=I("blue"))
```


##5. Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE, results='hide', include=TRUE}
d <- read.csv("activity.csv", header = TRUE,  na.strings = c(" ",NA))
na.omit(d)

d$days <- ifelse(isWeekday(d$date) == TRUE, 1, 0)
d
summary(d)

weekday <-d %>% filter(!is.na(steps)) %>% filter (days==1)
weekdayAvg <-na.omit(weekday %>% group_by(interval)) %>% summarise(totSteps = sum(steps))
weekend <- d %>% filter(!is.na(steps)) %>% filter (days==0)
weekendAvg <-na.omit(weekend %>% group_by(interval)) %>% summarise(totSteps = sum(steps))
summary(weekdayAvg)
```

``` {r echo=TRUE, include=TRUE, results='asis'}
ggplot() + 
  geom_line(data = weekdayAvg,  aes(x = interval, y = totSteps), color = "red") + 
  geom_line(data = weekendAvg,  aes(x = interval, y = totSteps), color = "blue") +
  labs(title="Histogram of avgerage number of steps over 5-minute interval")+ labs(subtitle="weekdays (red) vs. weekends (blue)")+
  xlab('5-min Intervals') +
  ylab('Avg. Steps')
```




